PyQBench: a Python library for benchmarking gate-based quantum computers

### 2023-02-11 txt/manuscript.tex
cli added

\begin{table}[H]
\begin{tabular}{|l|p{6.5cm}|p{6.5cm}|}
\hline
\textbf{Nr.} & \textbf{Code metadata description} & \textbf{Please fill in this
column} \\
\hline
C1 & Current code version & 0.1.0 \\
\hline
C2 & Permanent link to code/repository used for this code version & \url{https://github.com/iitis/PyQBench} \\
\hline
C3 & Code Ocean compute capsule & \todo[inline]{???}\\
\hline
C4 & Legal Code License & Apache License 2.0\\
\hline
C5 & Code versioning system used & git \\
\hline
C6 & Software code languages, tools, and services used & Python, Qiskit, AWS Braket \\
\hline
C7 & Compilation requirements, operating environments \& dependencies &
\texttt{Python >= 3.8}\newline
\texttt{numpy \textapprox= 1.22.0}\newline
\texttt{scipy \textapprox= 1.7.0}\newline
\texttt{pandas \textapprox= 1.5.0}\newline
\texttt{amazon-braket-sdk >= 1.11.1}\newline
\texttt{pydantic \textapprox= 1.9.1}\newline
\texttt{qiskit \textapprox= 0.37.2}\newline
\texttt{mthree \textapprox= 1.1.0}\newline
\texttt{tqdm \textapprox= 4.64.1}\newline
\texttt{pyyaml \textapprox= 6.0}\newline
\texttt{qiskit-braket-provider \textapprox= 0.0.3}\\
\hline
C8 & If available Link to developer documentation/manual &
\url{https://pyqbench.readthedocs.io/en/latest/}\\
\hline
C9 & Support email for questions & \url{dexter2206@gmail.com}\\
\hline
\end{tabular}
\caption{Code metadata (mandatory)}
\label{}
\end{table}

@@

\section{Preliminaries and discrimination scheme approach}
This section is divided into two parts.
In Section~\ref{sec:maths}, we introduce the necessary notation and preliminaries needed to understand the concept of PyQBench. Next, in Section~\ref{sec:discrimination-scheme},
we will present a general overview of the discrimination scheme in theoretical terms. We will also mention NISQ devices' limitations and possible practical solutions in the implementation process.

\subsection{Mathematical preliminaries}\label{sec:maths}
Recall that a general quantum
measurement, that is a positive operator valued measure (POVM) $\PP$ is a
collection of positive semidefinite operators $\{E_1, \ldots, E_m \}$ called
\emph{effects}, which sum up to identity, \ie $ \, \, \sum_{i=1}^m E_i = \1$.
In PyQBench, we are interested only in von Neumann measurements, i.e. measurements
for which all the effects are rank-one projectors. Every such measurement can be
parameterized by a unitary matrix $U$ which the effects $\{\proj{u_0}, \ldots, \proj{u_{d-1}}\}$,
are created by taking $\ket{u_i}$ as  $i+1$-th column of the unitary matrix $U$.
We will denote von Neumann measurements described by the matrix $U$ by $\PP_{U}$.

@@

\begin{figure}[h!]
	\centering
	\includegraphics[scale=1.7]{pics/vonneuman}
	\caption{Decomposition of von Neumann measurement}
	\label{fig:vonnneuman}
\end{figure}

@@

Without loss of generality, we consider discrimination between a measurement in the computational
Z-basis ($\PP_\Id$), and an alternative measurement performed in the basis $U$
($\PP_U$)\footnote{Explaining why we can consider only Z-basis and alternative measurement is beyond
	the scope of this technical documentation. See \cite{puchala2018strategies} if you are interested in
	the explanation.}. In PyQBench, we operate only on two-level systems, but the discrimination scheme,
described in detail in \cite{puchala2018strategies}, makes no assumptions about the dimensionality.

In general, the discrimination scheme, presented in Fig.\ref{fig:theoretical_scheme}, requires a
second system of the same dimensionality as the measured one. First, the joint system is prepared in
some state $\ket{\psi_0}$. Then, the unknown measurement is performed on the first part of the
system. Based on its outcome $i$, another measurement $\mathcal{P}_{V_i}$ is performed to obtain an
outcome $j$. Finally, if $j=0$ we guess that the performed measurement is $\mathcal{P}_U$, otherwise
we guess that it was $\mathcal{P}_\Id$.

@@

	The first idea is very simple. Instead of performing a conditional measurement, for each
	measurement $\PP_U, \PP_\Id$ to be discriminated and each choice of $k \in \{0, 1\}$ we run circuit presented in Fig. \ref{fig:postselection}.

@@

		\caption{
			Postselection scheme of discrimination between von Neumann measurements $\PP_{U}$ and $\PP_\Id$.
		}\label{fig:postselection}
	\end{figure}

	But now our experiment does not match the previously described scheme, right?
	Yes, unless we discard all the outcomes for which $i\ne k$ (hence the name \emph{postselection},
	we are selecting valid outcomes after the experiment is performed).

@@

	% \begin{enumerate}
	% \item We prepare  the discriminator $\ket{\psi_{0}}$ on bipartite system.
	% \item We perform one of two quantum channels, either $\Phi_{\Id}$ or
	% $\Phi_{U^\dagger}$,  on  the first part of the input state  $\ket{\psi_{0}}$, that means $\left( \Phi \otimes \Id\right)(\proj{\psi_0})$.
	% \item We measure the first part of the input state  $\ket{\psi_{0}}$ and
	% receive an output label $i$.
	% \item
	% Based on Holevo--Helstrom theorem, we perform a conditional binary
	% measurement	$\PP_{V_j}$ and hence, we implement the quantum channel $\Phi_{V_j^\dagger}$ on the second part.
	% \item To calculate the probability of correct discrimination, we include just those cases for which $i = j$.
	% \item We measure the second part of the input state  $\ket{\psi_{0}}$ and make a
	% decision based on received label $j$. If $i=j=0$, then we decide that
	% $\PP_U$ occurs.  If $i=j=1$, we decide that
	% $\PP_{\Id}$ occurs.
	% \end{enumerate}

	More precisely, our experiments can be grouped into classes identified by tuples of the form
	$(\mathcal{Q}, k, i, j)$, where $\mathcal{Q} \in \{\PP_U, \PP_\Id\}$ denotes the chosen measurement.
	We discard all the experiments for which $k \ne i$. Hence, the total number of valid experiments is

@@

	We now need to count the experiments (among the valid ones) resulting in successful discrimination.
	If we have chosen $\PP_U$, then we guess correctly iff $j=0$. Similarly, for
	$P_\Id$, we guess correctly iff $j=1$. If we define
	\begin{eqnarray}
	N_{\PP_U} &= \#\{(\mathcal{Q}, k, i, j): \mathcal{Q} = \PP_U, k = i, j = 0\}, \\
	N_{\PP_\Id} &= \#\{(\mathcal{Q}, k, i, j): \mathcal{Q} = \PP_\Id, k = i, j = 1\},
	\end{eqnarray}
	then the empirical success probability can be computed as

@@

	\begin{enumerate}
		\item We prepare the discriminator $\ket{\psi_{0}}$ on bipartite
		system.
		%\item We perform one of two quantum channels, either $\Phi_{\Id}$ or
		%$\Phi_{U^\dagger}$,  on  the first part of the input state  $\ket{\psi_{0}}$, that means $\left(\Phi \otimes \Id\right)(\proj{\psi_0})$.
		\item We apply one of unitary,   either $U^\dagger$ or
		$\Id$, on the first system.
		%\item Based on Holevo--Helstrom theorem, we perform a conditional binary measurement $\PP_{V_j}$, and hence, we implement the direct sum of quantum channels $\Phi_{V_0^\dagger} \oplus \Phi_{V_1^\dagger}$. Observe that   $\Phi_{V_0^\dagger} \oplus \Phi_{V_1^\dagger} = \proj{0} \otimes \Phi_{V_0^\dagger}  + \proj{1} \otimes \Phi_{V_1^\dagger}.  $
		\item We apply direct sum $V_0^\dagger \oplus V_1^\dagger$ on the whole systems.
		\item We measure both systems in computational basis $\Delta$.
		\item We make a decision based on the received label $j$ on the second system. If $j=0$, then we
		decide that $\PP_U$ occurs. Otherwise, we decide that $\PP_{\Id}$ occurs.
	\end{enumerate}

	The schematic representation of this setup is depicted in
	Fig.~\ref{fig:controlled}.
	\begin{figure}[h!]
		\centering
		\includegraphics[scale=1.5]{pics/controlled_unitary}

		\caption{ A schematic representation of the setup for distinguishing
			measurements using controlled unitary gate.
		}\label{fig:controlled}
	\end{figure}

@@

@@

\subsection{Software Architecture}\label{sec:sortware-architecture}
PyQBench can be used in two different ways:
\begin{itemize}
\item As a library. This is a more extensible, but also more laborious way of using PyQBench. Most importantly, this way allows for more advanced scripting defining your own measurements scheme.

\item As a CLI tool. This mode of operation trades extensibility for simplicity of usage, as it allows only for running Fourier-discrimination experiments defined with simple YAML input files.
\end{itemize}

\subsubsection{PyQBench as a library}\label{sec:pyqbench-as-library}
This guide introduces basic concepts needed for using PyQBench as a library from a Python script. In particular, we will cover the following topics:
\begin{enumerate}
\item Defining experiment using circuit components.
\item Assembling circuits needed for given benchmarking scheme.
\item Running experiment on simulator or hardware.
\item Obtaining empirical probability of successful discrimination between measurements.
\end{enumerate}

On the conceptual level, every discrimination experiment performed in PyQBench needs the following:
\begin{enumerate}
\item Discriminator, i.e. optimal initial state for the circuit.
\item Unitary $U^\dagger$ defining alternative measurement to be discriminated from the measurement in Z-basis.
\item $V_0^\dagger$ and $V_1^\dagger$, positive and negative parts of optimal measurement from Holevo-Helstrom theorem.
\end{enumerate}
Currently available quantum computers typically cannot start execution from an arbitrary state. Instead, we have to prepare it ourselves. Hence, to execute experiment using postselection scheme, we need an instruction taking to the desired discriminator
and instructions implementing $U^\dagger$, $V_0^\dagger$ and $V_1^\dagger$.

\subsubsection{Example of usage}
In our example, we will use $U =  H$ (the Hadamard gate). To keep us focused on the implementation in PyQBench and not delve too deep into mathematical explanation, we simply provide explicit formulas for discriminator $\ket{\Psi_0} $, $V_0$ and $V_1$, leaving the calculations to the interested reader.

The explicit formula for discriminator in our toy model reads:
\begin{equation}
\ket{\Psi_0} = \frac{1}{\sqrt{2}} (\ket{00} + \ket{11}),
\end{equation}
with corresponding parts of optimal measurement being equal to
\begin{equation}
V_0 =
\left(\begin{array}{cc} \alpha & -\beta\\ \beta & \alpha \end{array}\right),
\end{equation}
and \begin{equation}
V_1 =
\left(\begin{array}{cc} -\beta & \alpha \\ \alpha & \beta \end{array}\right),
\end{equation}
where \begin{equation}
\alpha = \frac{\sqrt{2 - \sqrt{2}}}{2} \cos\left( \frac{3}{8} \pi \right),
\end{equation}
\begin{equation}
\beta  = \frac{\sqrt{2  + \sqrt{2}}}{2} \sin\left( \frac{3}{8} \pi \right).
\end{equation}

As a next step, we need decompose our matrices into actual sequences of gates.
We are lucky, because our discriminator is just a Bell state. Thus, the circuit taking $\ket{00}$ to $\ket{\Phi_0}$ is well known, and comprises Hadamard gate followed by CNOT gate on both qubits.
\begin{figure}[h!]
	\centering
	\includegraphics[scale=1.7]{pics/discriminator}
	\caption{Decomposition of the discriminator $\ket{\Phi_0}$. }
	\label{fig:discriminator}
\end{figure}

For $V_0$ and $V_1$ observe that $V_0 = RY\left(\frac{3}{4} \pi \right) $, where $RY$ is rotation gate around the $Y$ axis defined by
\begin{equation}
RY(\theta) =
\left(\begin{array}{cc} \cos\frac{\theta}{2} & -\sin\frac{\theta}{2} \\ \sin\frac{\theta}{2} & \cos\frac{\theta}{2} \end{array}\right).
\end{equation}
To obtain $V_1$  we need only to swap the columns, which is equivalent to following $V_0$ by $X$ matrix, more precisely
\begin{equation}
V_1 =  RY\left(\frac{3}{4} \pi \right) \cdot X.
\end{equation}

Now running the simulation is as simple as invoking functions imported from qbench package.

\todo[inline]{listingi swieca sie na razie jak choinka ale bedziemy kolorki drasowac}
\begin{lstlisting}[language=Python, caption=Simulation benchmark by using postselection]

simulator = Aer.get_backend("aer_simulator")
postselection_result = benchmark_using_postselection(
	backend=simulator,
	target=0,
	ancilla=1,
	state_preparation=state_prep(),
	u_dag=u_dag(),
	v0_dag=v0_dag(),
	v1_dag=v1_dag(),
	num_shots_per_measurement=10000,
)
\end{lstlisting}

\begin{lstlisting}[language=Python, caption=Simulation benchmark by using direct sum]

simulator = Aer.get_backend("aer_simulator")
direct_sum_result = benchmark_using_direct_sum(
backend=simulator,
target=1,
ancilla=2,
state_preparation=state_prep(),
u_dag=u_dag(),
v0_v1_direct_sum_dag=v0_v1_direct_sum_dag(),
num_shots_per_measurement=10000,
)
\end{lstlisting}

\begin{lstlisting}[language=Python, caption=Computation of probability]
p_succ = (2 + np.sqrt(2)) / 4
print(f"Real p_succ = {p_succ}")
print(
f"Postselection: p_succ = {postselection_result}, abs. error = {p_succ - postselection_result}"
)
print(f"Direct sum: p_succ = {direct_sum_result}, abs. error = {p_succ - direct_sum_result}")

@@

\begin{lstlisting}
Real p_succ = 0.8535533905932737
Postselection: p_succ = 0.8559797193791593, abs. error = -0.0024263287858855564
Direct sum: p_succ = 0.85605, abs. error = -0.0024966094067262468
\end{lstlisting}

\subsubsection{Gaining more control over the benchmarking process}
In the example presented above we used functions that automate the whole process - from the circuit assembly, through running the simulations to interpreting the results. But what if we want more control over some parts of this process? Maybe we want play around with some models? Or maybe we want to run the same experiment on multiple backends and we don’t want them to be assembled over and over again?

 PyQBench provides functions performing:
\begin{enumerate}
\item assembly of circuits needed for experiment, provided the components discussed above
\item interpretation of the obtained measurements
\end{enumerate}
When using these functions, you need to take care of running circuits on some backend yourself, but it gives you unlimited power over the whole process. The difference between such approaches is illustrated on the diagrams below.

When using these functions, you need to take care of running circuits on some backend yourself, but it gives you unlimited power over the whole process. The difference between the two approaches is illustrated on the diagrams below.

\begin{figure}
	\centering
	\begin{subfigure}[b]{\textwidth}
		\centering
		\includegraphics[width=1\textwidth]{pics/scheme1}
%		\caption{$y=x$}
		\label{fig:simplified}
	\end{subfigure}
	\vfill
	\begin{subfigure}[b]{\textwidth}
		\centering
	\includegraphics[width=1\textwidth]{pics/scheme2}
%		\caption{$y=3\sin x$}
		\label{fig:execution}
	\end{subfigure}
\end{figure}

@@

Next, we divide the experiment into two parts:
\begin{enumerate}
	\item Assembling  circuits
	\item Running  circuits
\end{enumerate}

Let us focus only on the postselection case, as the direct sum case is analogous. In addition, we need to import two more functions from PyQBench.

\begin{lstlisting}[language=Python, caption=Assembling circuits]
from qbench.schemes.postselection import (
assemble_postselection_circuits,
compute_probabilities_from_postselection_measurements,
)

circuits = assemble_postselection_circuits(
target=0,
ancilla=1,
state_preparation=state_prep(),
u_dag=u_dag(),
v0_dag=v0_dag(),
v1_dag=v1_dag(),
)

circuits
\end{lstlisting}

%
%\begin{lstlisting}[language=Python, caption=Python example]
%{'id_v0': <qiskit.circuit.quantumcircuit.QuantumCircuit at 0x7f8b1585b9a0>,
%'id_v1': <qiskit.circuit.quantumcircuit.QuantumCircuit at 0x7f8b37fbfa90>,
%'u_v0': <qiskit.circuit.quantumcircuit.QuantumCircuit at 0x7f8b1585bf40>,
%'u_v1': <qiskit.circuit.quantumcircuit.QuantumCircuit at 0x7f8b1585bf10>}
%\end{lstlisting}

>>>>

\begin{table}[H]
\begin{tabular}{|l|p{6.5cm}|p{6.5cm}|}
\hline
\textbf{Nr.} & \textbf{Code metadata description} & \textbf{Please fill in this
column} \\
\hline
C1 & Current code version & 0.1.0 \\
\hline
C2 & Permanent link to code/repository used for this code version & \url{https://github.com/iitis/PyQBench} \\
\hline
C3 & Code Ocean compute capsule & \todo[inline]{???}\\
\hline
C4 & Legal Code License & Apache License 2.0\\
\hline
C5 & Code versioning system used & git \\
\hline
C6 & Software code languages, tools, and services used & Python, Qiskit, AWS Braket \\
\hline
C7 & Compilation requirements, operating environments \& dependencies &
\texttt{Python >= 3.8}\newline
\texttt{numpy \textapprox= 1.22.0}\newline
\texttt{scipy \textapprox= 1.7.0}\newline
\texttt{pandas \textapprox= 1.5.0}\newline
\texttt{amazon-braket-sdk >= 1.11.1}\newline
\texttt{pydantic \textapprox= 1.9.1}\newline
\texttt{qiskit \textapprox= 0.37.2}\newline
\texttt{mthree \textapprox= 1.1.0}\newline
\texttt{tqdm \textapprox= 4.64.1}\newline
\texttt{pyyaml \textapprox= 6.0}\newline
\texttt{qiskit-braket-provider \textapprox= 0.0.3}\\
\hline
C8 & If available Link to developer documentation/manual &
\url{https://pyqbench.readthedocs.io/en/latest/}\\
\hline
C9 & Support email for questions & \url{dexter2206@gmail.com}\\
\hline
\end{tabular}
\caption{Code metadata}
\label{}
\end{table}

@@

\section{Preliminaries and discrimination scheme approach}
This section is divided into two parts.
In Section~\ref{sec:maths}, we introduce the necessary notation and preliminaries needed to understand the concept of PyQBench. Next, in Section~\ref{sec:discrimination-scheme},
we will present a general overview of the discrimination scheme in theoretical terms. We will also mention NISQ devices' limitations and possible practical solutions during implementation process.

\subsection{Mathematical preliminaries}\label{sec:maths}
A general quantum
measurement, that is a positive operator valued measure (POVM) $\PP$ is a
collection of positive semidefinite operators $\{E_1, \ldots, E_m \}$ called
\emph{effects}, which sum up to identity, \ie $ \, \, \sum_{i=1}^m E_i = \1$.
In PyQBench, we are interested only in von Neumann measurements, i.e. measurements
for which all the effects are rank-one projectors. Every such measurement can be
parameterized by a unitary matrix $U$ which effects $\{\proj{u_0}, \ldots, \proj{u_{d-1}}\}$,
are created by taking $\ket{u_i}$ as  $(i+1)$-th column of the unitary matrix $U$.
We will denote von Neumann measurements described by the matrix $U$ by $\PP_{U}$.

@@

\begin{figure}[h!]
	\centering
	\includegraphics[scale=1.7]{pics/vonneuman}
	\caption{Decomposition of a von Neumann measurement $\PP_{U}$. }
	\label{fig:vonnneuman}
\end{figure}

@@

Without loss of generality, we consider discrimination task between a measurement in the computational
Z-basis ($\PP_\Id$), and an alternative measurement performed in the basis $U$
($\PP_U$)\footnote{Explaining why we can consider only Z-basis and alternative measurement $\PP_{U}$ is beyond
	the scope of this technical documentation. See \cite{puchala2018strategies} if you are interested in
	the explanation.}. In PyQBench, we operate only on two-level systems, but the discrimination scheme,
described in detail in \cite{puchala2018strategies}, makes no assumptions about the dimensionality.

In general, the discrimination scheme, presented in Fig.~\ref{fig:theoretical_scheme}, requires an
auxiliary system of the same dimensionality as the measured one. First, the joint system is prepared in
some state $\ket{\psi_0}$. Then, the unknown measurement is performed on the first part of the
system. Based on its outcome $i$, another measurement $\mathcal{P}_{V_i}$ is performed to obtain an
outcome $j$. Finally, if $j=0$ we guess that the performed measurement is $\mathcal{P}_U$, otherwise
we guess that it was $\mathcal{P}_\Id$.

@@

	The first idea uses a postselection. Instead of performing a conditional measurement, for each
	measurement $\PP_U, \PP_\Id$ to be discriminated and each choice of $k \in \{0, 1\}$ we run circuit presented in Fig. \ref{fig:postselection}.

@@

		\caption{
			A schematic representation of the setup for distinguishing
			measurements $\PP_{U}$ and $\PP_{\Id}$ using direct sum.
		}\label{fig:postselection}
	\end{figure}

%	But now our experiment does not match the previously described scheme, right?
%	Yes, unless we discard all the outcomes for which $i\ne k$ (hence the name \emph{postselection},
%	we are selecting valid outcomes after the experiment is performed).

@@

	% \begin{enumerate}
	% \item We prepare  the discriminator $\ket{\psi_{0}}$ on bipartite system.
	% \item We perform one of two quantum channels, either $\Phi_{\Id}$ or
	% $\Phi_{U^\dagger}$,  on  the first part of the input state  $\ket{\psi_{0}}$, that means $\left( \Phi \otimes \Id\right)(\proj{\psi_0})$.
	% \item We measure the first part of the input state  $\ket{\psi_{0}}$ and
	% receive an output label $i$.
	% \item
	% Based on Holevo--Helstrom theorem, we perform a conditional binary
	% measurement	$\PP_{V_j}$ and hence, we implement the quantum channel $\Phi_{V_j^\dagger}$ on the second part.
	% \item To calculate the probability of correct discrimination, we include just those cases for which $i = j$.
	% \item We measure the second part of the input state  $\ket{\psi_{0}}$ and make a
	% decision based on received label $j$. If $i=j=0$, then we decide that
	% $\PP_U$ occurs.  If $i=j=1$, we decide that
	% $\PP_{\Id}$ occurs.
	% \end{enumerate}
The experiments can be grouped into classes identified by tuples of the form
	$(\mathcal{Q}, k, i, j)$, where $\mathcal{Q} \in \{\PP_U, \PP_\Id\}$ denotes the chosen measurement, $k \in \{0,1\}$, $i \in \{0,1\}$ and $j \in \{0,1\}$.
	In postselection approach, we discard all the experiments for which $k \ne i$. Hence, the total number of valid experiments is

@@

	Finally, we count the experiments (among the valid ones) resulting in successful discrimination.
	If we have chosen $\PP_U$, then we guess correctly iff $j=0$. Similarly, for
	$P_\Id$, we guess correctly iff $j=1$. If we define
	\begin{eqnarray}
	N_{\PP_U} &= \#\{(\mathcal{Q}, k, i, j): \mathcal{Q} = \PP_U, k = i, j = 0\}, \\
	N_{\PP_\Id} &= \#\{(\mathcal{Q}, k, i, j): \mathcal{Q} = \PP_\Id, k = i, j = 1\},
	\end{eqnarray}
	then the empirical success probability can be computed as

@@

The second idea uses  the direct sum $V_0^\dagger \oplus V_1^\dagger.$    Here, instead of performing a conditional
measurement $\PP_{V_k}$, where $k\in \{0,1\}$ we run circuit presented in	Fig.~\ref{fig:controlled}.
%	\begin{enumerate}
%		\item We prepare the discriminator $\ket{\psi_{0}}$ on bipartite
%		system.
%		%\item We perform one of two quantum channels, either $\Phi_{\Id}$ or
%		%$\Phi_{U^\dagger}$,  on  the first part of the input state  $\ket{\psi_{0}}$, that means $\left(\Phi \otimes \Id\right)(\proj{\psi_0})$.
%		\item We apply one of unitary,   either $U^\dagger$ or
%		$\Id$, on the first system.
%		%\item Based on Holevo--Helstrom theorem, we perform a conditional binary measurement $\PP_{V_j}$, and hence, we implement the direct sum of quantum channels $\Phi_{V_0^\dagger} \oplus \Phi_{V_1^\dagger}$. Observe that   $\Phi_{V_0^\dagger} \oplus \Phi_{V_1^\dagger} = \proj{0} \otimes \Phi_{V_0^\dagger}  + \proj{1} \otimes \Phi_{V_1^\dagger}.  $
%		\item We apply direct sum $V_0^\dagger \oplus V_1^\dagger$ on the whole systems.
%		\item We measure both systems in computational basis $\Delta$.
%		\item We make a decision based on the received label $j$ on the second system. If $j=0$, then we
%		decide that $\PP_U$ occurs. Otherwise, we decide that $\PP_{\Id}$ occurs.
%	\end{enumerate}

%	The schematic representation of this setup is depicted in
%	Fig.~\ref{fig:controlled}.
	\begin{figure}[h!]
		\centering
		\includegraphics[scale=1.5]{pics/controlled_unitary}

		\caption{ A schematic representation of the setup for distinguishing
			measurements $\PP_{U}$ and $\PP_{\Id}$ using direct sum.
		}\label{fig:controlled}
	\end{figure}

@@

So far, we only discussed how the discrimination is performed using two different
schemes, assuming that all needed components ($\ket{\psi_0}$, $V_0$, $V_1$) are known. It turned out that the determination of $V_0$ and $V_1$ is a simple task while the discriminator $\ket{\psi_{0}}$ is known. However,
sad to say, determining the analytical form of the discriminator $\ket{\psi_{0}}$
is, in general, NP-hard problem.
We have found the analytical solution of the discrimination scheme for a particular class of von Neumann measurements.

@@

\subsection{Software Architecture}\label{sec:sortware-architecture}
PyQBench can be used in two different ways:
\begin{itemize}
\item As a library. This way allows for more advanced scripting defining your own measurements scheme.

\item As a CLI tool. This mode of operation trades extensibility for simplicity of usage, as it allows only for running Fourier-discrimination experiments defined with  YAML input files.
\end{itemize}

\section{PyQBench as a library}\label{sec:pyqbench-as-library}
This guide introduces basic concepts needed for using PyQBench as a library from a Python script. In particular, we will cover the following topics:
\begin{enumerate}
\item Defining experiment using circuit components.
\item Assembling circuits needed for given benchmarking scheme.
\item Running experiment on a simulator or a hardware.
\item Obtaining empirical probability of successful discrimination between measurements.
\end{enumerate}

On the conceptual level, every discrimination experiment performed in PyQBench needs the following:
\begin{enumerate}
\item Discriminator, i.e. optimal initial state for the circuit.
\item Unitary $U^\dagger$ defining alternative measurement to be discriminated from the measurement in Z-basis.
\item $V_0^\dagger$ and $V_1^\dagger$, positive and negative parts of optimal measurement from Holevo-Helstrom theorem.
\end{enumerate}
%Currently available quantum computers typically cannot start execution from an arbitrary state. Instead, we have to prepare it ourselves. Hence, to execute experiment using postselection scheme, we need an instruction taking to the desired discriminator $\ket{\psi_0}$
%and instructions implementing $U^\dagger$, $V_0^\dagger$ and $V_1^\dagger$. Whereas, execute experiment using direct sum approach, we need an instruction taking to the desired discriminator  $\ket{\psi_0}$
%and instructions implementing $U^\dagger$, $V_0^\dagger \oplus V_1^\dagger$.

\subsubsection{Example of usage}
In our example, we will use $U =  H$ (the Hadamard gate).
% To keep us focused on the implementation in PyQBench and not delve too deep into mathematical explanation, we simply provide explicit formulas for discriminator $\ket{\Psi_0} $, $V_0$ and $V_1$, leaving the calculations to the interested reader.
The explicit formula for discriminator in this model reads
\begin{equation}
\ket{\psi_0} = \frac{1}{\sqrt{2}} (\ket{00} + \ket{11}),
\end{equation}
with corresponding parts of optimal measurement being equal to
\begin{equation}
V_0 =
\left(\begin{array}{cc} \alpha & -\beta\\ \beta & \alpha \end{array}\right),
\end{equation}
and \begin{equation}
V_1 =
\left(\begin{array}{cc} -\beta & \alpha \\ \alpha & \beta \end{array}\right),
\end{equation}
where \begin{equation}
\alpha = \frac{\sqrt{2 - \sqrt{2}}}{2} \cos\left( \frac{3}{8} \pi \right),
\end{equation}
\begin{equation}
\beta  = \frac{\sqrt{2  + \sqrt{2}}}{2} \sin\left( \frac{3}{8} \pi \right).
\end{equation}
As a next step, we need decompose our matrices into actual sequences of gates.
The circuit taking $\ket{00}$ to the Bell state $\ket{\psi_0}$ is well known, and comprises Hadamard gate followed by CNOT gate on both qubits (see Fig.~\ref{fig:discriminator}).
\begin{figure}[h!]
	\centering
	\includegraphics[scale=1.7]{pics/discriminator}
	\caption{Decomposition of the discriminator $\ket{\psi_0}$. }
	\label{fig:discriminator}
\end{figure}
For $V_0$ and $V_1$ observe that $V_0 = RY\left(\frac{3}{4} \pi \right) $, where $RY$ is rotation gate around the $Y$ axis defined by
\begin{equation}
RY(\theta) =
\left(\begin{array}{cc} \cos\frac{\theta}{2} & -\sin\frac{\theta}{2} \\ \sin\frac{\theta}{2} & \cos\frac{\theta}{2} \end{array}\right).
\end{equation}
To obtain $V_1$  we need only to swap the columns, which is equivalent to following $V_0$ by $X$ matrix, more precisely
\begin{equation}
V_1 =  RY\left(\frac{3}{4} \pi \right) \cdot X.
\end{equation}
Finally, the optimal probability of correct discrimination between $\PP_{U}$, where $U=H$ and $\PP_{\Id}$ is given by 
\begin{equation}
p_{\text{succ}}(\PP_{U}, \PP_{\Id}) = \frac{1}{2} + \frac{\sqrt{2}}{4}. 
\end{equation}
The detailed calculation we can see in \ref{app:hadamard}.

Now we simulate  theoretical framework using PyQBench. We will start with simulation in Qiskit SDK with Aer simulator. First, we import necessary modules. \todo[inline]{czy to zdanie jest poprawne?} 
\todo[inline]{listingi swieca sie na razie jak choinka ale bedziemy kolorki drasowac}

\begin{lstlisting}[language=Python]
from qiskit import QuantumCircuit, Aer
import numpy as np
\end{lstlisting}
Next we import needed functions from PyQBench. For our example we will need two functions. 
\begin{lstlisting}[language=Python]
from qbench.schemes.postselection import benchmark_using_postselection
from qbench.schemes.direct_sum import benchmark_using_direct_sum
\end{lstlisting}
To run any of these functions, we need to define components that we discussed below. Most importantly, we need to create Qiskit instructions, not circuits. We can conveniently do so by constructing circuit acting on qubits 0 and 1 and then converting them using \textit{to\_instruction()} method.
\begin{lstlisting}[language=Python, caption= Defining components for Hadamard experiment]
def state_prep():
	circuit = QuantumCircuit(2)
	circuit.h(0)
	circuit.cnot(0, 1)
	return circuit.to_instruction()

def u_dag():
	circuit = QuantumCircuit(1)
	circuit.h(0)
	return circuit.to_instruction()

def v0_dag():
	circuit = QuantumCircuit(1)
	circuit.ry(-np.pi * 3 / 4, 0)
	return circuit.to_instruction()

def v1_dag():
	circuit = QuantumCircuit(1)
	circuit.ry(-np.pi * 3 / 4, 0)
	circuit.x(0)
	return circuit.to_instruction()

def v0_v1_direct_sum_dag():
	circuit = QuantumCircuit(2)
	circuit.ry(-np.pi * 3 / 4, 0)
	circuit.cnot(0, 1)
	return circuit.to_instruction()

\end{lstlisting}
For the purpose of this example,  we define Qiskit Aer simulator.
\begin{lstlisting}[language=Python, caption= Defining a backend]
simulator = Aer.get_backend("aer_simulator")

\end{lstlisting}
Now running the simulation is as simple as invoking functions imported from PyQBench package.
\begin{lstlisting}[language=Python, caption=Simulation benchmark by using postselection]
postselection_result = benchmark_using_postselection(
	backend=simulator,
	target=0,
	ancilla=1,
	state_preparation=state_prep(),
	u_dag=u_dag(),
	v0_dag=v0_dag(),
	v1_dag=v1_dag(),
	num_shots_per_measurement=10000,
)
\end{lstlisting}

\begin{lstlisting}[language=Python, caption=Simulation benchmark by using direct sum]
direct_sum_result = benchmark_using_direct_sum(
	backend=simulator,
	target=1,
	ancilla=2,
	state_preparation=state_prep(),
	u_dag=u_dag(),
	v0_v1_direct_sum_dag=v0_v1_direct_sum_dag(),
	num_shots_per_measurement=10000,
)
\end{lstlisting}
Finally, we calculate the empirical probability obtained from postselection and direct sum experiment and calculate the absolute error between analytical and empirical results. 
\begin{lstlisting}[language=Python, caption=Computation of probability]
p_succ = (2 + np.sqrt(2)) / 4
print(f"Analytical p_succ = {p_succ}")
print(
f"Postselection: p_succ = {postselection_result}, abs. error = {p_succ - postselection_result}"
)
print(f"Direct sum: p_succ = {direct_sum_result}, abs. error = {p_succ - direct_sum_result}")

@@

\begin{lstlisting}
Analytical p_succ = 0.8535533905932737
Postselection: p_succ = 0.8559797193791593, abs. error = -0.0024263287858855564
Direct sum: p_succ = 0.85605, abs. error = -0.0024966094067262468
\end{lstlisting}

\subsubsection{Gaining more control over the benchmarking process}
In the example presented above we used functions that automate the whole process - from the circuit assembly, through running the simulations to interpreting the results. But what if we want more control over some parts of this process? 
One possibility would be to add more and more parameters to \textit{benchmark\_using\_xyz} functions, but this approach clearly is not very scalable. Also, there is no way we can anticipate all the possible use cases.
 We decided on another approach.  PyQBench provides functions performing:
\begin{enumerate}
\item assembly of circuits needed for experiment, provided the components discussed above
\item interpretation of the obtained measurements
\end{enumerate}
 The difference between the two approaches is illustrated on the diagrams below.

\begin{figure}[ht!]
	\centering
	\begin{subfigure}[b]{\textwidth}
		\centering
		\includegraphics[width=0.8\textwidth]{pics/scheme1}
%		\caption{$y=x$}
		\label{fig:simplified}
	\end{subfigure}
	\vfill
	\begin{subfigure}[b]{\textwidth}
		\centering
	\includegraphics[width=0.8\textwidth]{pics/scheme2}
		\label{fig:execution}
	\end{subfigure}
\end{figure}

@@

By choosing this approach we need to divide the experiment into two parts:
\begin{enumerate}
	\item Assembling  circuits
	\item Running  circuits
\end{enumerate}
Let us focus only on the postselection case, as the direct sum case is analogous. In addition, we need to import two more functions from PyQBench.
\begin{lstlisting}[language=Python, caption=Assembling circuits]
from qbench.schemes.postselection import (
	assemble_postselection_circuits,
	compute_probabilities_from_postselection_measurements,
)
\end{lstlisting}
The function \textit{assemble\_postselection\_circuits} creates all of neceserry circuits, nicely packed in a dictionary. Each informs if the circuit implements identity or alternative measurement on the target qubit, and which of $V_0$ and
$V_1$ is used.
\begin{lstlisting}[language=Python]
circuits = assemble_postselection_circuits(
	target=0,
	ancilla=1,
	state_preparation=state_prep(),
	u_dag=u_dag(),
	v0_dag=v0_dag(),
	v1_dag=v1_dag(),
)
